---
title: "Holdout"
output: html_document
date: "2024-04-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages

The tidyverse is loaded in for you in the code chunk below. The visualization package, ggplot2, and the data manipulation package, dplyr, are part of the “larger” tidyverse.

```{r load_packages}
library(tidyverse)
```
The modelr package is loaded in for you in the code chunk below. You may use functions from modelr to calculate performance metrics for your models.

```{r load_packages1}
library(modelr)
```

The caret package to manage all aspects of data splitting, training, and evaluating the models.

```{r load_packages2}
library(caret)
```

## Import Data

```{r read_data}
# Importing training data
train_data_path <- 'paint_project_train_data.csv'
df_train<- readr::read_csv(train_data_path, col_names = TRUE)
```

```{r read_holdout_data}
# Importing training data
holdout_data_path <- 'paint_project_holdout_data.csv'
df_holdout<- readr::read_csv(holdout_data_path, col_names = TRUE)
```

```{r glimpse}
df_holdout %>% glimpse()
```
```{r LogitTransform}
logit <- function(p) {
  log(p / (1 - p))
}

df_train$logit_response <- logit((df_train$response - 0) / (100 - 0))
```

## Holdout Test Set Requirements



```{r holdout}

fit_lm <- lm(logit_response ~ R + G + B + Hue + Lightness + Saturation, data = df_train)
fit_glm <- glm(outcome ~ Lightness + Saturation + R + G + B + Hue , family = binomial, data = df_train)
df_holdout$y <- predict(fit_lm, newdata = df_holdout, type = "response")
df_holdout$probability <- predict(fit_glm, newdata = df_holdout, type = "response")
df_holdout$outcome <- ifelse(df_holdout$probability > 0.5, "event", "non_event")
df_predictions <- df_holdout %>%
  rowid_to_column("id") %>%
  select(id, y, outcome, probability)
write_csv(df_predictions, "C:/Users/harsh/OneDrive/Desktop/Spring 2024/Machine Learning/Harshitha_Venkatesh_Project_HAV41/predictions.csv")




```

